UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-04-23 ms/main.tex
abstract

\begin{abstract}
{\large Abstract.}
%The abstract is typically 100-150 words, unreferenced. 
Improvements in sequencing have generated massive and diverse genomic datasets of populations. Dimension reduction is necessary to condense and analyze the data. Uniform manifold approximation and projection (UMAP) is a non-linear dimension reduction tool recently made available. Here we apply UMAP to three well-studied genotype datasets and discover overlooked subpopulations within the American Hispanic population, fine-scale relationships between geography, genotypes, and phenotypes in the UK population, and cryptic structure in the Thousand Genomes Project data. We demonstrate this approach outperforms other standard methods, is well-suited to the influx of large and diverse data, and opens new lines of inquiry in population-scale datasets.
\end{abstract}

>>>>

\begin{abstract}
{\large Abstract.}
%The abstract is typically 100-150 words, unreferenced. 
Human populations feature both discrete and continuous patterns of variation. Current analysis approaches struggle to jointly identify these patterns because of modelling assumptions, mathematical constraints, or numerical challenges. Here we apply uniform manifold approximation and projection (UMAP), a non-linear dimension reduction tool, to three well-studied genotype datasets and discover overlooked subpopulations within the American Hispanic population, fine-scale relationships between geography, genotypes, and phenotypes in the UK population, and cryptic structure in the Thousand Genomes Project data. We demonstrate this approach outperforms other standard methods, is well-suited to the influx of large and diverse data, and opens new lines of inquiry in population-scale datasets.
\end{abstract}
